Functions - Hadley master class
========================================================

## define data
```{r make-data-set-1}
set.seed(123)
df <- data.frame(x=round(rnorm(20, mean=99, sd=5)),
                 a=round(rnorm(20, mean=10, sd=10)),
                 b=round(rnorm(20, mean=0, sd=3)),
                 c=round(rnorm(20, mean=50, sd=2)))
df
```

## Apply DRY (Don't Repeat Yourself) principle
```{r fix-missing}
fix_missing <- function(x) {
  x[x == -99] <- NA
  x
}

df[] <- lapply(df, fix_missing)
```

```{r explore-1}
numeric <- vapply(df, is.numeric, logical(1))
numeric
df[numeric] <- lapply(df[numeric], fix_missing)
df
```

```{r get-rid-of-repetition}
summary <- function(x) {
  c(mean(x), median(x), sd(x), mad(x), IQR(x))
}

summary(df$a)
summary(df$b)
summary(df$c)
summary(df$x)

summary <- function(x) {
  c(mean(x, na.rm=TRUE), 
    median(x, na.rm=TRUE), 
    sd(x, na.rm=TRUE), 
    mad(x, na.rm=TRUE), 
    IQR(x, na.rm=TRUE))
}

summary(df$a)
summary(df$b)
summary(df$c)
summary(df$x)
```

## Examine parts of a function
```{r examine-parts-of-function}
formals(function(x=4) g(x) + h(x))
body(function(x) g(x) + h(x))
environment(function(x) g(x) + h(x))
```

## Closures
```{r closures-1}
x <- 5
f <- function() {
  y <- 10
  c(x=x, y=y)
}
f()
g <- function() {
  x <- 20
  y <- 10
  c(x=x, y=y)
}
g()
```

```{r closures-2}
x <- 0
y <- 10
f <- function(){
  x <- 1
  function() {
    y <- 2
    x + y
  }
}
```

Create a function that creates a whole class of functions
```{r closures-3}
power <- function(exponent) {
  function(x) x^exponent
}

square <- power(2)
c(square(2), square(3))
cube <- power(3)
c(cube(2), cube(3))

# examine square
# find environment and its parent
environment(square)
parent.env(environment(square))

# inspect objects defined in that environment
ls(environment(square))
get('exponent', environment(square))
environment(square)$exponent
as.list(environment(square))
```

Multiple counters
```{r multiple-counters}
new_counter <- function() {
  i <- 0
  function() {
    # do something useful, then...
    i <<- i + 1
    i
  }
}
counter_one <- new_counter()
counter_two <- new_counter()

counter_one()
counter_two()
counter_two()
counter_one()
```

Built in functions that make closures
```{r built-in-fn-make-closures}
Negate(is.numeric)('abc')
Negate

?Vectorize
vrep <- Vectorize(rep.int, 'times')
vrep(42, times=1:4)
vrep
as.list(environment(vrep))

e <- ecdf(runif(1000))
str(e)
e(0.5)
class(3)

```

## Higher Order Functions
```{r higher-order-functions}
# Data structure HOFs
# Provide basic tools for when you have a predicate
# function instead of a logical vector.
# Filter: keeps true
# Find: value of first true
# Position: location of first true
head(Filter(is.factor, iris))
head(Find(is.factor, iris))
Position(is.factor, iris)
# One function I use a lot:
# Not sure what this does - get rid of null values?
# null != NA
compact <- function(x) Filter(Negate(is.null), x)

set.seed(123)
samples <- replicate(5, sample(10, 20, rep = T),
  simplify = FALSE)
# Want to find intersection of all values
int <- intersect(samples[[1]], samples[[2]])
int <- intersect(int, samples[[3]])
int <- intersect(int, samples[[4]])
int <- intersect(int, samples[[5]])
```

Simple example of Reduce
```{r}
# Reduce recursively applies a function in this way
Reduce(intersect, samples)
add <- function(x, y) x + y
Reduce(add, 1:10)
Reduce(add, 1:10, 5)
Reduce(add, 1:10, accumulate=TRUE)
length(Reduce(add, 1:10, accumulate=TRUE))
```

## Mathematical HOFs
```{r math-hofs}
integrate(sin, 0, pi)
uniroot(sin, pi * c(1 / 2, 3 / 2))
optimise(sin, c(0, 2 * pi))
optimise(sin, c(0, 2 * pi))[[1]] / pi
optimise(sin, c(0, pi), maximum = TRUE)
optimise(sin, c(0, pi), maximum = TRUE)[[1]] / pi
```

## Combine closure and HOF
## MLE
```{r combine-closure-and-hof}
# Combination of closures and HOF particularly useful.
# For statistics, maximum likelihood estimation is a
# great example.
poisson_nll <- function(x) {
  n <- length(x)
  function(lambda) {
    n * lambda - sum(x) * log(lambda) # + ...
  }
}
nll1 <- poisson_nll(c(41, 30, 31, 38, 29, 24, 30, 29))
nll2 <- poisson_nll(c(6, 4, 7, 3, 3, 7, 5, 2, 2, 7))
optimise(nll1, c(0, 100))
optimise(nll2, c(0, 100))
```
 ## Lists of Functions
 
```{r lists-of-functions}
compute_mean <- list(
  base = function(x) mean(x),
  sum = function(x) sum(x) / length(x),
  manual = function(x) {
    total <- 0
    n <- length(x)
    for (i in seq_along(x)) {
      total <- total + x[i] / n
    }
total }
)
call_fun <- function(f, ...) f(...)
x <- runif(1e6)
lapply(compute_mean, call_fun, x)
lapply(compute_mean, function(f) system.time(f(x)))
```


### Modify summary to return return a user specified list of functions
```{r modify-summary}
summary <- function(x) {
 c(mean(x, na.rm = TRUE),
   median(x, na.rm = TRUE),
   sd(x, na.rm = TRUE),
   mad(x, na.rm = TRUE),
   IQR(x, na.rm = TRUE))
}
summary(df$a)
summary(df$b)
summary(df$c)

summary2 <- function(x, ys) {
  call_fun <- function(f, ...) f(...)
  lapply(ys, call_fun, x)
}
summary2(df$a, compute_mean)
summary2(1:100, list(
  length = function(x) length(x),
  sum = function(x) sum(x)))
```
