Variations of Box Plots in R for Ozone Concentrations in New York City and Ozonopolis
========================================================

http://chemicalstatistician.wordpress.com/2013/05/26/exploratory-data-analysis-variations-of-box-plots-in-r-for-ozone-concentrations-in-new-york-city-and-ozonopolis/

* Chambers, J. M., Cleveland, W. S., Kleiner, B. and Tukey, P. A. (1983) Graphical Methods for Data Analysis. Wadsworth & Brooks/Cole.
* Kabacoff, Robert. R in Action. Manning Publications Co., 2011.
* http://vita.had.co.nz/papers/boxplots.pdf

```{r get-data}
ozone <- airquality$Ozone
boxplot(ozone, ylab= 'Ozone (ppb)', main= 'Box Plot of Ozone in New York')
```

Interpret box plot
* top and bottom of box are 3rd and 1st quartiles
* line in rectangle is median
* top whisker is max value or 3rd quartile + 1.5 * IQR (IQR = Q3 - Q1) (whichever is smaller)
* bottom whisker is minimum value or 1st quartile - 1.5 * IQR (whichever is larger)

Summary
```{r summary}
summary(ozone)
fivenum(ozone)
```

## simulate ozone pollution data for ozonopolis
Create a 2nd set with rgamma()
```{r simulate-ozonopolis}
set.seed(1)
mean.ozone <- mean(ozone, na.rm=TRUE)
var.ozone <- var(ozone, na.rm=TRUE)
n <- sum(is.na(ozone))
ozone2 <- rgamma(n, shape=mean.ozone^2 / var.ozone + 3, scale=var.ozone / mean.ozone + 3)
```

plot both box plots
```{r plot-both-boxplots}
boxplot(ozone, ozone2, ylab='Ozone (ppb)', names=c('New York', 'Ozonopolis'), main='Box Plots for NY and Ozonopolis')

# Horizontal
boxplot(ozone, ozone2, xlab = 'Ozone (ppb)', names = c('New York', 'Ozonopolis'), main = 'Box Plots for New York and Ozonopolis', horizontal = T)
```

### Using axis() to specify the axes
```{r using-axis}
boxplot(ozone, ozone2, ylab = 'Ozone (ppb)', xaxt = 'n', main = 'Box Plots for New York and Ozonopolis')
axis(1, at = c(1, 2), labels = c('New York', 'Ozonopolis'))
```
* xaxt='n' suppresses default x-axis labels
* 1st argument of axis()
  * 1 = bottom
  * 2 = left
  * 3 = top
  * 4 = right


### Add notches
If notches don't overlap, the 2 medians are significantly different with 95% confidence
```{r plot-with-notches}
boxplot(ozone, ozone2, notch = T, ylab = 'Ozone (ppb)', names = c('New York', 'Ozonopolis'), main = 'Box Plots with Notches for New York and Ozonopolis', col = c('blue', 'red'))

```

