From Cookbook for R - Colors(ggplot2)
=============

## Install libraries
```{r libraries}
library(ggplot2)
library(RColorBrewer)
```

## Sample data
```{r sample-data}
df <- read.table(header=T, text='
cond yval
A 2
B 2.5
C 1.6
')

# Three variables
df2 <- read.table(header=T, text='
 cond1 cond2 yval
    A      I 2
    A      J 2.5
    A      K 1.6
    B      I 2.2
    B      J 2.4
    B      K 1.2
    C      I 1.7
    C      J 2.3
    C      K 1.9
')
```


## Simple color assignment
```{r simple-color-assignment}
# default
ggplot(df, aes(x=cond, y=yval)) + geom_bar()
# bars with red outlines
ggplot(df, aes(x=cond, y=yval)) + geom_bar(colour='#FF9999')
ggplot(df, aes(x=cond, y=yval)) + geom_bar(stat='identity', fill=rgb(1, 0, 0, .2))
# red fill, black outlines
ggplot(df, aes(x=cond, y=yval)) + geom_bar(fill='#FF9999', colour='black')

# Standard black lines and points
ggplot(df, aes(x=cond, y=yval)) + 
  geom_line(aes(group=1)) + # group points otherwise no line will show
  geom_point(size=3)

# Dark blue lines, red dots
ggplot(df, aes(x=cond, y=yval)) + 
  geom_line(aes(group=1), colour='#000099') +
  geom_point(size=3, colour='#cc0000')
```

## Map variable values to colors
```{r map-variable-to-color}
# bars: x and fill both depend on cond2
ggplot(df, aes(x=cond, y=yval, fill=cond)) + geom_bar(stat='identity')

# bars with other dataset, fill depends on cond2
ggplot(df2, aes(x=cond1, y=yval)) +
  geom_bar(aes(fill=cond2),
           colour='black',
           position=position_dodge())

ggplot(df2, aes(x=cond1, y=yval, colour=cond2)) +
  geom_line(aes(group=cond2)) +
  geom_point(size=3)
```

## Colorblind-friendly palette
Define your own palette for later use
```{r colorblind-friendly-palette}
# Can store palette to a variable then use it later - HOW?
# The palette with grey:
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# The palette with black:
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# To use for fills, add
scale_fill_manual(values=cbPalette)

# To use for line and point colors, add
scale_colour_manual(values=cbPalette)
```

## Color selection
```{r color-selection}
# equivalent - by default, scale_fill_hue() is used
ggplot(df, aes(x=cond, y=yval, fill=cond)) + geom_bar(stat='identity')
ggplot(df, aes(x=cond, y=yval, fill=cond)) + geom_bar(stat='identity') + scale_fill_hue()

ggplot(df, aes(x=cond, y=yval, colour=cond)) + geom_point(size=8)
ggplot(df, aes(x=cond, y=yval, colour=cond)) + geom_point(size=8) + scale_colour_hue()

```

## Change luminance and saturation (chromacity)
```{r change-luminance-and-saturation}
# Change luminance to 40 (instead of default 65)
ggplot(df, aes(x=cond, y=yval, fill=cond)) + geom_bar(stat='identity') + scale_fill_hue()
ggplot(df, aes(x=cond, y=yval, fill=cond)) + geom_bar(stat='identity') + scale_fill_hue(l=40)
# Change luminance to 40, and reduce saturation (chromacity) from 100 to 50
ggplot(df, aes(x=cond, y=yval, fill=cond)) + geom_bar(stat='identity') + scale_fill_hue()
ggplot(df, aes(x=cond, y=yval, fill=cond)) + geom_bar(stat='identity') + scale_fill_hue(c=45)
ggplot(df, aes(x=cond, y=yval, fill=cond)) + geom_bar(stat='identity') + scale_fill_hue(l=80, c=45)
```

## Using RColorBrewer
```{r using-brewer}
ggplot(df, aes(x=cond, y=yval, fill=cond)) + geom_bar(stat='identity') + scale_fill_brewer()
ggplot(df, aes(x=cond, y=yval, fill=cond)) + geom_bar(stat='identity') + scale_fill_brewer(palette='Set1')
ggplot(df, aes(x=cond, y=yval, fill=cond)) + geom_bar(stat='identity') + scale_fill_brewer(palette='Spectral')
# Note: use scale_colour_brewer() for lines and points
```

## Manually define Palette
Define your own set of colors with scale_fill_manual()
```{r manually-defined-palette}
ggplot(df, aes(x=cond, y=yval, fill=cond)) + geom_bar(stat='identity') + 
  scale_fill_manual(values=c('red', 'blue', 'green'))
ggplot(df, aes(x=cond, y=yval, fill=cond)) + geom_bar(stat='identity') + 
  scale_fill_manual(values=c('#cc6666', '#9999cc', '#66cc99'))
# Note: use scale_colour_manual() for lines and points
```

## Continuous colors
```{r continuous-colors}
set.seed(133)
df2 <- data.frame(xval=rnorm(50), yval=rnorm(50))

# make color depend on yval
ggplot(df2, aes(x=xval, y=yval, colour=yval)) + geom_point()

# use a different gradient
ggplot(df2, aes(x=xval, y=yval, colour=yval)) + geom_point() + scale_colour_gradientn(colours=rainbow(2))
ggplot(df2, aes(x=xval, y=yval, colour=yval)) + geom_point() + scale_colour_gradientn(colours=rainbow(4))
ggplot(df2, aes(x=xval, y=yval, colour=yval)) + geom_point() + scale_colour_gradientn(colours=rainbow(10))
```

Qualitative Brewer Packages good for qualitative data

Accent 8

Dark2 8

Paired 12

Pastel1 9

Pastel2 8

Set1 9

Set2 8

Set3 12

# Create a palette and display it
```{r create-display-palette}
pal <- brewer.pal(9, 'Greens')[c(3, 5, 7, 9)]
pal <- c(pal, brewer.pal(9, 'Blues')[c(3, 5, 7, 9)])
pal <- c(pal, brewer.pal(9, 'Reds')[c(3, 5, 7, 9)])
pal <- c(pal, brewer.pal(9, 'Purples')[c(3, 5, 7, 9)])
pal <- c(pal, brewer.pal(9, 'BrBG')[c(4, 3, 2, 1)])
pal <- c(pal, brewer.pal(9, 'Greys')[c(5, 6, 7, 8)])
image(1:30, 1, as.matrix(1:30), col=pal, xlab='Greens (sequential)', ylab='', xaxt='n', yaxt='n', bty='n')

# display a divergent palette
display.brewer.pal(7, 'BrBG')

# display a qualitative palette
display.brewer.pal(7, 'Accent')

# Display n palettes simultaneously
display.brewer.all(n=10, exact.n=FALSE)
display.brewer.all(n=10)
display.brewer.all()
display.brewer.all(type='div')
display.brewer.all(type='seq')
display.brewer.all(type='qual')
display.brewer.all(n=5, type='div', exact.n=TRUE)
brewer.pal.info

brewer.pal.info['Blues',]
brewer.pal.info['Blues',]$maxcolors
```

## Create a combined palette?
```{r create-combined-palette}
pal <- c(brewer.pal(n=12, name='Paired'), brewer.pal(n=9, name='Pastel1'), brewer.pal(n=9, name='Set1'))
pal <- unique(pal)
pal
```

Quick and dirty colors
```{r colors}
# pal <- c(brewer.pal(n=12, name='Paired'), brewer.pal(n=9, name='Pastel1'), brewer.pal(n=9, name='Set1'))
# pal <- unique(pal)
# pal
pal <- brewer.pal(9, 'Greens')[c(3, 5, 7, 9)]
pal <- c(pal, brewer.pal(9, 'Blues')[c(3, 5, 7, 9)])
pal <- c(pal, brewer.pal(9, 'Reds')[c(3, 5, 7, 9)])
pal <- c(pal, brewer.pal(9, 'Purples')[c(3, 5, 7, 9)])
pal <- c(pal, brewer.pal(9, 'BrBG')[c(4, 3, 2, 1)])
pal <- c(pal, brewer.pal(9, 'Greys')[c(4, 6, 7, 8)])
# image(1:length(pal), 1, as.matrix(1:length(pal)), col=pal, xlab='mypalette', ylab='', xaxt='n', yaxt='n', bty='n')
```