Data Prep
========================================================

From http://www.r-bloggers.com/data-preparation-part-i/?utm_source=feedburner&utm_medium=email&utm_campaign=Feed%3A+RBloggers+%28R+bloggers%29

## Using apply, lapply, tapply
```{r using-apply}
set.seed(123)
matriz <- matrix(round(runif(12, 1, 10), 0), nrow=3)
```

apply(X, MARGIN, FUN, ...)
MARGIN = 1 = row
MARGIN = 2 = column
MARGIN = c(1, 2) = rows and columns

```{r apply-2}
matriz
apply(matriz, 1, sum)  # Sum by row
apply(matriz, 2, sum)  # Sum by column
apply(matriz, c(1, 2), function(x) x^2)  # square everything
```

## tapply - useful on descriptive analysis
tapply(X, INDEX, FUN = NULL, ..., simplify = TRUE)
INDEX = list of one or more factors
```{r tapply-mtcars}
head(mtcars)
str(mtcars)
# mean power by cylinder capacity
tapply(mtcars$hp, mtcars$cyl, mean)
```

If you have a list, can use lapply or sapply (simplify the output)
```{r lapply-sapply}
# Generate data
lista <- list(a=c('one', 'tow', 'three'), b=c(1,2,3), c=c(12, 'a')) 
lista

# How many elements in each list?
lapply(lista, length)  # return a list
sapply(lista, length)  # return a vector
```

## Split - apply - combine without plyr
Use split, *apply, rbind/cbind

### Split
```{r split}
data <- split(mtcars, mtcars$gear)  # split
data
```

### Apply
```{r apply}
fits <- lapply(data, function(x) return(lm(x$mpg ~ x$disp)$coef))
fits
```

### Combine
```{r combine}
do.call(rbind, fits)
```


