Best distribution of blue and red balls
========================================================

FB interview question
1/16/14

- You have 100 marbles, 1/2 are red, and the other 1/2 are blue.
- Place the balls in 2 bags such that you maximize the probability of picking
a blue ball.
- Assume the probability of picking each bag is the same  
P (blue) = P (blue|bag1)P (bag1) + P (blue|bag2)P (bag2)  
= 1/2 × [P (blue|bag1) + P (blue|bag2)]  

So we want to maximize  
P (blue|bag1) + P (blue|bag2) 

Caution: Don’t assume you have the same number of balls in each bag! define  
b = number of blue balls in bag 1 (5) r = number of red balls in bag 1 (6)  
P (blue|bag1) = b / b+r  
P (blue|bag2) = (50 - b) / (50−b)+(50−r)  
```{r libs}
library(plyr)
library(reshape2)
library(ggplot2)
```

```{r prob-blue}
prob_b <- function(b, r) {
  b/(b + r) + (50 - b) / (100 - b - r)
}
```

```{r create-df}
df <- data.frame(r = 0:50)
df <- mutate(df, b0 = prob_b(0, r), b1 = prob_b(1, r), b5 = prob_b(5, r), b49 = prob_b(49, r), b50 = prob_b(50, r))
df.tidy <- melt(df, id=c('r'))
head(df.tidy)

df1 <- data.frame(r = 0:50)
dfb <- data.frame(b = c(0:5, seq(10, 25, 5)))
dfj <- expand.grid(df1$r, dfb$b)
dfj <- rename(dfj, replace=c('Var1' = 'r', 'Var2' = 'b'))
head(dfj)
dfj <- mutate(dfj, pb = prob_b(b, r))
```

```{r plot-it}
ggplot(df.tidy, aes(r, value, group=variable, color=variable)) + geom_line()
ggplot(dfj, aes(r, pb, group=b, color=b)) + geom_line()
ggplot(dfj, aes(r, b, color=pb)) + geom_point()
```

