Using gridExtra
========================================================
Source
* http://www.r-bloggers.com/extra-extra-get-your-gridextra/

```{r load-libraries-and-data}
library(ggplot2) #for plotting
library(reshape2) #for data reshaping
library(gridExtra)

library(vegan) #for the data
data(varechem)
```
 
First, reshape the data so that Hummus depth and Bare soil are your grouping variables
```{r}
vMelt<-melt(varechem, measure.vars=c("Humdepth", "Baresoil"))
```
 
Now plot it.  Use fill to color things differently, facet_wrap to split this into two panels,
And don't forget that the x scales are different - otherwise things look odd
```{r}
qplot(value, data=vMelt, fill=variable)+facet_wrap( facets=~variable, scale="free_x")
```

This produces a nice graph. But, man, I had to think about reshaping things, and all of those scales? What if I just wanted to make two historgrams, and slam ‘em together. This is where gridExtra is really nice. Through its function grid.arrange, you can make a multi-paneled graph using ggplot2 plots, lattice plots, and more (although, not regular R plots…I think).

So, let’s see the same example, but with gridExtra.

```{r side-by-side-2-plots}
library(gridExtra)
 
#make two separate ggplot2 objects
humDist<-qplot(Humdepth, data=varechem, fill=I("red"))
bareDist<-qplot(Baresoil, data=varechem, fill=I("blue"))
 
#Now use grid.arrange to put them all into one figure.
#Note the use of ncol to specify two columns.  Things are nicely flexible here.
grid.arrange(humDist, bareDist, ncol=2)
```

“Oh, what a trivial problem,” you may now be saying. But, if you want to, say, plot up 5 different correlations, or, say, the same scatterplot with 4 different model fits, this is a life-saver – if nothing else, in terms of readability of your code for later use.

This is all well and good, but, simple. Let’s get into more fun multi-panel figures. Let’s say we wanted a bivariate scatter-plot of Hummus Depth and Bare Soil with a linear fit. But, we also wanted to plot the histograms of each variable in adjacent panels. Oh, and flip the histogram of whatever is on the y-axis. Sexy, no? This is pretty straightforward. We can use the ggplot2 objects we already have, flip the co-ordinates on one, create a bivariate plot with a fit, and fill in one final panel with something blank.

```{r include-blank-plot}
#First, the correlation.  I'm using size just to make bigger points.  And then I'll add a smoothed fit.
corPlot<-qplot(Humdepth, Baresoil, data=varechem, size=I(3))+stat_smooth(method="lm")
 
#OK, we'll need a blank panel.  gridExtra can make all sorts of shapes, so, let's make a white box
blankPanel<-grid.rect(gp=gpar(col="white"))
 
#Now put it all together, but don't forget to flip the Baresoil histogram
grid.arrange(humDist, blankPanel, corPlot, bareDist, ncol=2)
```

Nice. Note the use of the grid.rect. gridExtra is loaded with all sorts of interesting ways to place shapes and other objects into your plots – including my favorite – grid.table, for when you don’t want to deal with text.

```{r grid-table}
a<-anova(lm(Baresoil ~ Humdepth, data=varechem))
grid.table(round(a, digits=3))
```

Or, heck, if you want to make that part of the above plot, use tableGrob instead of grid.table, and then slot it in where the blank panel is. The possibilities are pretty endless!
```{r tableGrob}
tbl <- tableGrob(round(a, digits=3))
grid.arrange(humDist, tbl, corPlot, bareDist, ncol=2)

```
