Argon - t-test
========================================================

http://chemicalstatistician.wordpress.com/2013/03/10/discovering-argon-with-the-2-sample-t-test/

##### Analyzing Lord Rayleigh's Data on Nitrogen and Discovering Argon with the 2-Sample t-Test
##### Written by Eric Cai - The Chemical Statistician

```{r define-argo-data}


atmospheric.nitrogen = c(2.31017, 2.30986, 2.31010, 2.31001, 2.31024, 2.31010, 2.31028, NA)
chemical.nitrogen = c(2.30143, 2.29890, 2.29816, 2.30182, 2.29869, 2.29940, 2.29849, 2.29869)
nitrogen.masses = data.frame(atmospheric.nitrogen, chemical.nitrogen)
colnames(nitrogen.masses) = c('Nitrogen from Air', 'Nitrogen from Chemical Decomposition')
```


```{r nitrogen-box-plot}
boxplot(nitrogen.masses, main = "Lord Rayleigh''s Measured Masses of Nitrogen", ylab = 'Mass (g)')

```

## Do a 2 sample t-test
* Null Hypothesis: There is no difference between the masses of the 2 types of nitrogen.
* Alternative Hypothesis: There is a difference between the masses of the 2 types of nitrogen.
```{r two-sample-t-test-argon}
t.test(atmospheric.nitrogen, chemical.nitrogen)
```

Interpretation
* Low p-value indicates less evidence for null hypothesis
* p-value < 0.05, so reject null hypothesis - so, 2 samples of nitrogen are different!